\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-14 metatheory-condensed.tex
Adding Inheritance Information.

$$
\Rule[name=Tm Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}},
\Rule[]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
\quad
\Rule[name=Dbj Indices]
{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}},
\Rule[]
{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
$$

% Explicit substitution
Instead of meta-level substitution, we use \textbf{explicit substitution}, i.e., substitution itself is a expression in the program and will be done in the runtime. Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$. We can apply substitution to type and terms in MLTT. We also have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
% Debruijn Indices
Contrary to named variables, \textbf{debruijn indices} refer to the variable in the context like operating a stack. For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{\pi_2 "id"}{T[\pi_1"id"]}$. Similarly, to get the second and above variables, we only need to apply a sequence of $\pi_1$s.

@@

The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$. For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators). Then any later fields $\goodTerm{\Gamma, A}{t}{T}$ relies only on the \textit{interface} $A$ can be inherited to other contexts (e.g. using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e. $\Gamma, A$)  as well.

@@

Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair $A \vdash B$ as well), but with mulitple constructors. Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$. Unconventionally, a signature can induce a "universe" of types $\bW$: in other words, $T : \bW$ will be the inductive type instead of $\bW$ itself. "Wsup" is used to construct a specific term of an given inductive type.

Similar to plugin implementation, we delegate handlers for recursion using linkages. We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers. Specifying the handler family using "Wrec" on target inductive term $w$ leads to elimination.

% Inhertance judgement
In the plugin implementation, "Family"s are second class objects like Coq's "Module". What's more, a "Family", with or without "extends" argument can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family. Thus we model them using \textit{inheritance judgement} instead of considering them as first class terms.

In this way, we will consider the program in our metatheory similar to those vanilla MLTT program, except there are meta variables denoting inheirtance judgement, of which the only usage is to apply to linkages. We don't yet consider non-trivial equality between inheritance judgements.

>>>>

$$
\Rule[name=Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}},
\Rule[]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
\quad
\Rule[name=Dbj Indices]
{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}},
\Rule[]
{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
$$

% Explicit substitution
Instead of meta-level substitution, we use \textbf{explicit substitution}, i.e., substitution itself is a expression in the program and will be done in the runtime. Thus substitution has its own judgement $\goodSub{\Gamma}{\gamma}{\Delta}$. We can apply substitution to types\footnote{The type judgement is the form $\goodType{\Gamma}{T}{}$} and terms\footnote{The term judgement is the form $\goodTerm{\Gamma}{t}{T}$} in MLTT via \ruleref{Subst}. We also have an identity substitution $\goodSub{\Gamma}{"id"}{\Gamma}$.
% Debruijn Indices
Contrary to named variables, \textbf{debruijn indices} refer to the variable in the context like operating a stack. For example, we can get the very first variable in the context $\Gamma, x : T \vdash x : T$ by $\goodTerm{\Gamma, T}{\pi_2 "id"}{T[\pi_1"id"]}$. Similarly, to get the second and above variables, we only need to apply a sequence of $\pi_1$s.

@@

The sealing judgement provides a witness $f$ that a context $\Gamma, \cC\sigma$ can be abstracted into an \textit{interface} $\Gamma, A$ for a type $\goodType{\Gamma}{A}{}$. For example, sealing judgement can abstract an inductive type with constructors into an arbitrary type with functions (but without eliminators). Then any later fields $\goodTerm{\Gamma, A}{t}{T}$ relies only on the \textit{interface} $A$ can be inherited to other contexts (e.g. using extended inductive type), as long as the targeting contexts can be abstracted into the surrounding \textit{interface} for $t$ (i.e. $\Gamma, A$)  as well.

@@

Our formulation of (extensible) inductive type is close to Wtype (\ruleref{WSig-Add} requires a pair $A \vdash B$ as well), but with mulitple constructors. Thus we have the signature of inductive type collecting each constructors, with a brand new judgement ${\goodWSig{\Gamma}{\tau}{n}}$. Unconventionally, a signature can induce a "universe" of types $\bW\tau$---in other words, $T : \bW\tau$ will be the inductive type instead of $\bW\tau$ itself. "Wsup" is used to construct a specific term of a given inductive type.

Similar to plugin implementation, we delegate handlers for recursion to linkages. We use "RecSig1" to compute the type of a single handler and "RecSig" for the complete family of handlers. Specifying the handler family using "Wrec" on target inductive term $w$ leads to elimination.

$$
\Rule[name=Inh]
{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
\quad \goodTerm{\Gamma}{l}{\cL \sigma_1}
}
{\goodTerm{\Gamma}{("inh" \ h \ l)}{\cL \sigma_2}},
\Rule[name=Inh-Id]
{}
{\goodInh{\Gamma}{"inhid"}{\sigma}{\sigma}},
\Rule[name=Inh-Ext]
{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}
  \quad \goodTerm{\Gamma, A_2}{t}{T}}
{\goodInh{\Gamma}{"inhext" \ h \ t}{\sigma_1}{(\nu^+ \  \sigma_2\  T)}}
$$

% Inhertance judgement
In the plugin implementation, "Family"s are second class objects like Coq's "Module". What's more, a "Family" (with or without "extends" clause) can be considered as a standalone piece of code trying to \textit{inherit} an existent (but maybe empty) family. Thus we model them using \textit{inheritance judgement} $\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}$ instead of considering them as first class terms. The only way to use $h$ is by \ruleref{Inh} and transforming a linkage of signature $\sigma_1$ to another linkage of $\sigma_2$.

Inheritance judgement can be further enriched with inheriting operation, overriding operation, extending operation and etc. These operations correspond to inherit, override and extend command inside "Family" during plugin implementation.

In this way, we will consider the program in our metatheory similar to those vanilla MLTT program, except there are meta variables denoting inheirtance judgement (i.e. "Family"), of which the only usage is to apply \ruleref{Inh}. We don't yet consider non-trivial equality between inheritance judgements.
